### 2018-12-07 README.Rmd
initial README development

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# areal

While methods to interpolate multiple spatial data sets are currently available in the ArcGIS Suite and the `R` package `sf`, `areal` offers an intuitive assortment of options to validate, interpolate, and compare fields between a source and target data set.

## Installation

The best way to install the released version of areal is from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("areal")
```

The development version of `areal` can be accessed from Github with `devtools`:

``` r
# install.packages("devtools")
devtools::install_github("charlie-revord/areal")
```

## Usage

`areal` is split into several different utilities that include: validation of source and target data  (specifically tests for sf object status, shared area unit type, and shared coordinate system), unit type conversion, intersection of source and target data, source field weight calculation, and target field calculation based off source field weight.

## Basic Use

The following showcases the capabilities of `areal` given that the source and target data

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# areal
While methods to interpolate multiple spatial data sets are currently available in the ArcGIS Suite and the `R` package `sf`, `areal` offers an intuitive assortment of options to validate, interpolate, and compare fields between a source and target data set.

## Installation
The development version of `areal` can be accessed from Github with `remotes`:

``` r
# install.packages("remotes")
remotes::install_github("slu-openGIS/areal")
```

## Usage 
The package contains two overlapping data sets, `stlRace` (2017 ACS demographic counts at the census tract level; $n=106$) and `stlWards` (the 2010 political subdivisions in St. Louis; $n=28$). These can be used to illustrate the core functionality of the package.

The basic usage of `areal` is through the `aw_interpolate()` function. One advantage of `areal` is that it allows for multiple values to be interpolated at a time. Here, both total estiatmed population (`TOTAL_E`) and the estimated count of white residents (`WHITE_E`) are interpolated at once:

```r
> library(areal)
>
> race <- stlRace
> wards <- stlWards
>
> aw_interpolate(wards, tid = WARD, source = race, sid = "GEOID", output = "tibble", "TOTAL_E", "WHITE_E")
# A tibble: 28 x 4
    WARD       AREA TOTAL_E WHITE_E
 * <int>      <dbl>   <dbl>   <dbl>
 1     1  46138761.   7992.    153.
 2     2 267817711.  12145.   1323.
 3     3  66291644.   7344.    591.
 4     4  53210707.   8458.    160.
 5     5  60462396.   8783.   1526.
 6     6  64337271.  14050.   5840.
 7     7 101268146.  15840.   8220.
 8     8  45966410.  12188.   7604.
 9     9  73993891.  14217.   6838.
10    10  62915358.  11239.   8703.
# ... with 18 more rows
```

One key advantage of `areal` is that is pipeable, meaning that it can fit easily into existing `tidyverse` workflows:

```r
> wards %>%
+ select(-OBJECTID, -AREA) %>%
+ aw_interpolate(tid = WARD, source = race, sid = "GEOID", output = "tibble", "TOTAL_E", "WHITE_E")
# A tibble: 28 x 3
    WARD TOTAL_E WHITE_E
 * <int>   <dbl>   <dbl>
 1     1   7992.    153.
 2     2  12145.   1323.
 3     3   7344.    591.
 4     4   8458.    160.
 5     5   8783.   1526.
 6     6  14050.   5840.
 7     7  15840.   8220.
 8     8  12188.   7604.
 9     9  14217.   6838.
10    10  11239.   8703.
# ... with 18 more rows
```

Another advantage of `areal` is that the interpolation process is not a "black box", but rather can be manually completed if necessary. Details on this are available on the package's website.
